---
title: "PA 1"
output: html_document
---

## 1) Loading and preprocessing data

```{r}
url <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url,"activity.zip")
activity <- read.table( unzip("activity.zip"),colClasses=c("numeric","character","numeric"),sep=",",header=TRUE)
colnames(activity)
str(activity)
nrow(activity)
sum(is.na(activity$steps))
head(activity$steps)

```

## 2) Mean total number of steps taken per day

```{r}
data <- activity[complete.cases(activity),]
data$date <- as.factor(data$date)
head(data)
str(data)
steps.sum <- tapply(data$steps,data$date,sum)
length(steps.sum)
head(steps.sum)
hist(steps.sum,breaks=10)
rmean <- mean(steps.sum)
rmedian <- median(steps.sum)
```
The mean is `r rmean` and the median is `r rmedian`.

## 3) Daily average activity pattern

```{r}
daily.average <- tapply(data$steps,data$interval,mean)
plot(daily.average,type="l")
max(daily.average)
m <- match(max(daily.average),daily.average)
m
k <- data$interval[m]
k
```
The maximum is the `r m`-th interval, at minute `r k`

## 4) Imputting missing values

```{r}
sum(!complete.cases(activity))
data.fill <- activity
data.fill$date <- as.factor(data.fill$date)
data.fill[is.na(data.fill$steps),"steps"] <- daily.average[match(data.fill$interval[is.na(data.fill$steps)],as.numeric(levels(as.factor(data.fill$interval))))]
str(data.fill)
fill.sum <- tapply(data.fill$steps,data.fill$date,sum)
hist(fill.sum,breaks=20)
hist(fill.sum,breaks=10)
mean(fill.sum)
median(fill.sum)
```
The median differs from the original histogram. 
By filling in NA values regardless of the method, creates data out of nothing, and thus introduces bias.

## 5) Weekday vs. Weedend activity patterns

```{r}
data.fill$date <- as.POSIXct(data.fill$date, format="%Y-%m-%d")
data.fill$date <- weekdays(data.fill$date,abbreviate=TRUE)
week <- function(x){
        y <- vector(mode="character",length(x))
        for (i in 1:length(x)){
                if (x[i] == "S" | x[i] == "D") {y[i] <- "weekend"
                }
                else {y[i] <- "weekday"
                }
        }
        y
}
data.week <- cbind(data.fill,as.factor(week(data.fill$date)))
colnames(data.week)[4] <- "weekday"
str(split(data.week,data.week$weekday))
final <- lapply(split(data.week,data.week$weekday),function(x){
        tapply(x$steps,x$interval,mean)
})
str(final)
final.fantasy <- as.data.frame(final)
bins <- as.numeric(rownames(table(activity$interval)))
l <- length(table(activity$interval))
final.fantasy$weekend <- as.numeric(final.fantasy$weekend)
final.fantasy$weekday <- as.numeric(final.fantasy$weekday)
final.final <- as.data.frame(cbind(c(final.fantasy$weekday,final.fantasy$weekend),c(rep("weekday",l),rep("weekend",l)),c(bins,bins)))
colnames(final.final) <- c("step","day","int")
final.final$step <- as.numeric(as.character((final.final$step)))
final.final$int <- as.numeric(as.character((final.final$int)))
str(final.final)
library(lattice)
xyplot(step~int|day,data=final.final,type="l")
```

During weekdays there is a peak in activity in the morning, wich is lacking in weekends. 
Also in the weekend the activity level seems slightly higher than on weekdays.
(seems like honest working people waking up in the morning, and then socializing in the weekends)